#Reproducible Research: Peer-graded Assignment: Course project 1
================================================================

##Explicitely setting the defaults and loading R packages


```{r setoptions, echo=TRUE, results="asis"}

library(ggplot2)
library (dplyr)
library(lattice)

```

##1.Loading the data

It is assumed that the Working Directory is set correctly, i.e. is set to where the file "repdata_data_activity.zip" is saved. 

```{r Question1}

unzip(zipfile="repdata_data_activity.zip")

data <- read.csv("activity.csv")

```

##2. & 3. Histogram of the total number of steps taken each day, and mean and median

```{r Question2-3}

dayly.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)

qplot(dayly.steps, binwidth=1000, xlab="Total Number of Steps taken each Day")

mean(dayly.steps, na.rm=TRUE)

median(dayly.steps, na.rm=TRUE)

```

##4. Time series plot of the average number of steps taken

```{r Question4}

average.steps <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
ggplot(data=average.steps, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-Minute Interval") +
    ylab("Average Number of Steps taken")

```

##5. The 5-minute interval that, on average, contains the maximum number of steps

```{r Question5}

average.steps[which.max(average.steps$steps),]

```

##6. Code to describe and show a strategy for imputing missing data


This is the percentage of intervals that have no values for the number of steps in that interval:

```{r}

mean(is.na(data$steps))*100

```

This is the total number of intervals that have no values for the number of steps in that interval:

```{r}

sum(is.na(data$steps))

```

We will replace each missing value with the mean.

```{r}

cleaned.data<-data

for (row.num in 1:nrow(cleaned.data)) {

    if (is.na(cleaned.data[row.num, "steps"])) {

        interval.sgn<- cleaned.data[row.num, "interval"]
        interval.sgn.ind <- which(average.steps[, "interval"] == interval.sgn)
        interval.steps.mean <- average.steps[interval.sgn.ind, "steps"]
        cleaned.data[row.num, "steps"] <- interval.steps.mean
         }
}

```

##7. Histogram of the total number of steps taken each day after missing values are imputed

```{r Question7}

steps.cleaned <- tapply(cleaned.data$steps, cleaned.data$date, FUN=sum)

qplot(steps.cleaned, binwidth=1000, xlab="Total Number of steps per Day")

mean(steps.cleaned)
median(steps.cleaned)

```
##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r Question8}

Sys.setenv(LANGUAGE = "en")
Sys.setlocale("LC_TIME", "English")

days.of.week <- weekdays(as.Date(cleaned.data$date))

cleaned.data$day.type <- sapply(days.of.week, function(which.day) {
    if (which.day %in% c("Saturday", "Sunday")) {
        return("weekend")
    } else {
        return("weekday")
           }
    })

steps.per.daytype <- cleaned.data %>% group_by(interval, day.type) %>% summarise(avg.steps = mean(steps))

xyplot(avg.steps ~ interval | day.type, data = steps.per.daytype, type = "b", layout = c(1, 2))

```